### 2017-03-03 doc/seminar2017.Rmd
Included birth control in table 1

---
title: "Intro Seminar"
author: "Windy Wang"
date: "March 1, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, echo = FALSE, message = FALSE, warning = FALSE)
library(dplyr)
devtools::load_all()
load_data(update = TRUE)
set_options()
```

```{r dsSetup}
ds <- project_data
# 
# dsBase <- ds %>% 
#   dplyr::filter(fVN == "Baseline")
# 
# dsComplete <- ds %>% 
#   dplyr::group_by(SID) %>%
#   dplyr::filter(n() == 3) %>%
#   dplyr::ungroup()
```

Subject Characterization
========================

Subject characteristics across visit numbers
```{r subject characteristics}

subCharTable <- tableone::CreateTableOne(
  vars = c("Age",
           "Sex",
           "Ethnicity",
           "BMI",
           "Waist",
           "eGFR",
           "ACR",
           "UrineCreatinine",
           "UrineMicroalbumin",
           "UrinaryCalcium",
           "UDBP",
           "Creatinine",
           "VitaminD",
           "PTH",
           "ALT",
           "Systolic",
           "Diastolic",
           "MeanArtPressure",
           "Glucose0",
           "Glucose120",
           "dmStatus",
           "SmokeCigs",
           "BirthControl"),
  strata = c("fVN"),
  data = ds,
  factorVars = c("Sex", "SmokeCigs", "BirthControl")
) %>% 
  print(nonnormal = c("Glucose0",
                      "Glucose120",
                      "ALT",
                      "UDBP",
                      "ACR")) %>% 
  knitr::kable()
```

Part 1 - Cross-Sectional
========================

## ACR

```{r acr}

# Box plot of uVDBP in different albuminuria categories
ds %>% 
  filter(fVN == "Baseline") %>% 
  select(acrStatus, udbpCrRatio) %>%
  na.omit() %>%
  box_plot_slides("acrStatus", "log(udbpCrRatio)", 
            "Albuminuria",
            "log uVDBP:Creatinine")

# ANOVA
anova <- aov(formula = log(udbpCrRatio)~acr_status, data = ds1)
summary(anova)
TukeyHSD(anova)
rm(anova)
```

```{r acr scatterplot}

# Scatterplot of ACR and uVDBP ----------------------------------

ds %>% 
  filter(fVN == "Baseline") %>% 
  select(ACR, udbpCrRatio) %>% 
  na.omit() %>% 
  scatter_plot("log(ACR)", "log(udbpCrRatio)",
               "log Albumin:Creatinine Ratio",
               "log UDBP:Creatinine",
               line = TRUE)

# Spearman Correlation ------------------------------------------

ds %>% 
  filter(fVN == "Baseline") %>% 
  cor.test(formula = ~ ACR + udbpCrRatio, data = ., method = "spearman")
```

## eGFR

```{r eGFR boxplot}

# Boxplot of uVDBP concentrations across eGFR categories --------------

ds %>% 
  filter(fVN == "Baseline") %>% 
  select(eGFRStatus, udbpCrRatio) %>% 
  na.omit() %>% 
  box_plot_slides("eGFRStatus", "log(udbpCrRatio)", 
            "Estimated GFR (ml/min/1.73m^2)",
            "log uVDBP:Creatinine")

# ANOVA

anova <- aov(formula = log(udbpCrRatio)~eGFR_status, data = ds1)
summary(anova)
TukeyHSD(anova)
rm(anova)
```

```{r eGFR scatterplot}

# Scatterplot of eGFR and uVDBP ----------------------------------

ds %>% 
  dplyr::filter(!(eGFR_status == "Moderate")) %>% 
  dplyr::filter(eGFR < 150) %>% 
  dplyr::select(eGFR, udbpCrRatio) %>% 
  na.omit() %>% 
  scatter_plot("eGFR", "log(udbpCrRatio)",
               "Estimated Glomerular Filtration Rate (ml/min/1.73m^2)",
               "log UDBP:Creatinine") +
  ggplot2::geom_smooth(method = lm, colour = "grey50")

# Spearman Correlation ------------------------------------------

ds %>% 
  filter(!(acr_status == "Macroalbuminuria")) %>% 
  cor.test(formula = ~ eGFR + udbpCrRatio, data = ., method = "spearman")
```

```{r hyperfiltration}
ds %>% 
  dplyr::filter(eGFR < 45) %>% 
  dplyr::select(SID, VN, Age, eGFR, dm_status, acr_status)

ds %>% 
  scatter_plot("Age", "eGFR",
               "Age", "eGFR")
```

## Glycemic Status

```{r DM boxplot}

# Boxplot of uVDBP concentrations across DM categories

ds %>% 
  filter(fVN == "Baseline") %>% 
  select(dmStatus, udbpCrRatio) %>% 
  na.omit() %>% 
  box_plot_slides("dmStatus", "log(udbpCrRatio)", 
            "Glycemic Status",
            "log uVDBP:Creatinine")

# ANOVA

anova <- aov(formula = log(udbpCrRatio)~eGFR_status, data = ds1)
summary(anova)
TukeyHSD(anova)
rm(anova)
```

Part 2 - GEE
============

```{r correlation}
dsGEE <- ds %>% 
  dplyr::mutate(udbpBase = ifelse(fVN == "Baseline", UDBP, NA),
                ageBase = ifelse(fVN == "Baseline", Age, NA),
                bmiBase = ifelse(fVN == "Baseline", BMI, NA),
                waistBase = ifelse(fVN == "Baseline", Waist, NA), 
                Sex = ifelse(Sex == "Male", 1, 0),
                Ethnicity = ifelse(Ethnicity == "European", Ethnicity, "Other"),
                Ethnicity = ifelse(Ethnicity == "European", 0, 1),
                fMedsBP = ifelse(fMedsBP == "Yes", 1, 0)
                # vitdBase = ifelse(fVN == "6Year", NA, VitaminD)
                ) %>% 
  dplyr::arrange(SID, fVN) %>% 
  dplyr::group_by(SID) %>% 
  tidyr::fill(udbpBase, ageBase, bmiBase, waistBase) %>% 
  dplyr::ungroup() %>%
  dplyr::arrange(SID, VN) %>% 
  dplyr::select(udbpBase, ACR, eGFR, VN, ageBase, Sex, Ethnicity, BMI, 
                Waist, DM, fMedsBP)

round(cor(dsGEE[sapply(dsGEE, is.numeric)], use = "complete.obs"), 2)

# Heatmap

corrplot::corrplot(cor(dsGEE[sapply(dsGEE, is.numeric)], use = "complete.obs"),
                   method = "color")

```

```{r mason}

gee <- ds %>%
  dplyr::mutate(
  udbpBase = ifelse(fVN == "Baseline", UDBP, NA),
  ageBase = ifelse(fVN == "Baseline", Age, NA),
  fDM = relevel(as.factor(DM), "1")
  # Ethnicity = ifelse(Ethnicity == "European", Ethnicity, "Other"),
  ) %>%
  dplyr::filter(!(fVN == "Baseline" &
                    acr_status == "Macroalbuminuria")) %>%
  dplyr::filter(!(fVN == "Baseline" & eGFR_status == "Moderate")) %>%
  dplyr::arrange(SID, fVN) %>%
  dplyr::group_by(SID) %>%
  tidyr::fill(udbpBase, ageBase) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(UDBP = UDBP/1000) %>%
  dplyr::arrange(SID, VN) %>%
  mason_gee(
  yvars = c("ACR", "eGFR"),
  xvars = "UDBP",
  covars = c("VN", "ageBase", "fDM")
  )
```

```{r}
gee3 <- readRDS(file = "../data/ds.Rds") %>%
  dplyr::mutate(
  udbpBase = ifelse(fVN == "Baseline", UDBP, NA),
  ageBase = ifelse(fVN == "Baseline", Age, NA),
  VitaminD3 = ifelse(fVN == "6Year", NA, VitaminD)
  # Ethnicity = ifelse(Ethnicity == "European", Ethnicity, "Other"),
  ) %>%
  dplyr::arrange(SID, fVN) %>%
  dplyr::group_by(SID) %>%
  tidyr::fill(udbpBase, ageBase) %>%
  dplyr::ungroup() %>%
  # dplyr::group_by(VN) %>%
  # dplyr::mutate(udbpBase = as.numeric(scale(udbpBase))) %>%
  dplyr::mutate(UDBP = UDBP/1000) %>%
  # dplyr::ungroup() %>% 
  dplyr::arrange(SID, VN) %>%
  mason_gee(
  yvars = c("VitaminD3"),
  xvars = "UDBP",
  covars = c("VN", "ageBase", "DM")
  )
```

Graph
```{r gee plot}

gee %>%
  dplyr::mutate(p.value = ifelse(p.value > 0.05, 1, 0.04)) %>%
  seer::view_main_effect(
  graph.options = 'dot.size',
  groups = '~Yterms',
  legend.title = 'p-values',
  xlab = 'Percent difference with 95% CI in the outcomes\nfor each unit increase in baseline uVDBP',
  ylab = 'Kidney Measures'
  ) +
  graph_theme(ticks = FALSE, legend.pos = 'right') +
  ggplot2::theme(legend.margin = grid::unit(0, 'cm'))

```